---
title: "Binding strength prediction"
author: "Richel J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This demo shows how to predict the binding strength
of peptides to different Major Histocompatibility
MHC haplotypes. 

First load the library:

```{r}
library(netMHCpan)
```

This demo assumes NetMHCpan is installed.
Installation cannot be done `netMHCpan`, as one needs
to request a download link at 
[http://www.cbs.dtu.dk/cgi-bin/nph-sw_request?netMHCpan/](http://www.cbs.dtu.dk/cgi-bin/nph-sw_request?netMHCpan).

To install `netMHCpan`, use `install_netMHCpan` with the download link:

```
install_netMHCpan(" ")
```

The installation of `netMHCpan` is checked, with
the goal of producing a helpful error message:

```{r}
tryCatch(
  check_netMHCpan_installation(),
  error = function(e) print(e)
)
```

Now, lets use the sequence of an example protein:

```{r}
sequence <- "FAMILYVWFAMILYV"
message(sequence)
```

Now, we need to select an MHC haplotype. There are many alleles,
so first we count the number of haplotypes:

```{r}
if (is_netMHCpan_installed()) {
  mhc_haplotypes <- get_netMHCpan_alleles()
  length(mhc_haplotypes)
}
```

Now, we simply pick the first haplotype:

```{r}
if (is_netMHCpan_installed()) {
  mhc_haplotype <- mhc_haplotypes[1]
}
```

Now, we can predict how strong our peptide binds to our allele, by obtaining the Inhibitory Concentration 50% (IC50) 
value in nanomolars (nM), of which lower values indicate stronger
binders:

```{r}
if (is_netMHCpan_installed()) {
  knitr::kable(
    predict_ic50(
      peptides = sequence,
      mhc_haplotype = mhc_haplotype
    )
  )
  
}
```
To investigate if whole a protein is immunogenic, we need to
obtain the IC50 values for all its cleaved products.
As the MHC-II molecules prefers 13 amino acids residues,
we can get the IC50 values for each residue as such:

```{r}
if (is_netMHCpan_installed()) {
  knitr::kable(
    predict_ic50s(
      protein_sequence = "AVLWAGVAFLAFLQLTALVLNLL",
      peptide_length = 13,
      mhc_haplotype = mhc_haplotype
    )
  )
  
}
```

